Integration of national demographic-disturbance relationships into local caribou population viability projections can reduce uncertainty and inform monitoring decisions

### 2025-01-24 analysis/paper/paper.Rmd
Revisions in progress.

resultScenario = "S10"
resultSet ="cmult6ay1aSf1"
figSet = c("diffs","distScns","EVSI","power","EVPIall")

@@

Examining the probability of
correctly assessing population status (Figure \@ref(fig:figStatusErrors))
and the expected value of information (Figures \@ref(fig:figEVSI) and E.1) helps clarify how the need for monitoring varies with landscape condition.
When anthropogenic disturbance is high (right panels of Figures
\@ref(fig:figStatusErrors), \@ref(fig:figEVSI), and E.1)), projections
of population status informed only by priors from the national
demographic-disturbance relationships are accurate, and adding
additional local monitoring data does not improve accuracy.
Interestingly, when the projection period is long and anthropogenic
disturbance is high, we see a negative EVSI (lower right panel of Figure
\@ref(fig:figEVSI)), but we do not see an associated decrease in the
probability of correctly assessing population status (lower right panel
of Figure \@ref(fig:figStatusErrors)). This means that the long-term
projections informed by local monitoring data are less
precise but not less accurate, which can happen if the prior viability predictions are more certain than the posterior predictions (e.g 
left side of right panels in Figure \@ref(fig:examplePredictions)). At
lower levels of disturbance, local monitoring can reduce uncertainty, but
there are diminishing returns as the number of years of monitoring increases (Figures
\@ref(fig:figStatusErrors) and \@ref(fig:figEVSI)). Reducing uncertainty is more difficult when interannual variation is high (Figures
\@ref(fig:figStatusErrors) and \@ref(fig:figEVSI)).The shapes of the
relationships between monitoring effort and the probability of correct
status assessment depend on how anthropogenic disturbance is changing
over time, but there are diminishing benefits of increased monitoring
regardless of whether disturbance is increasing or decreasing
(Supplement D). Acknowledging the possibility of composition survey errors increases uncertainty, but these increases are not large for the set of parameters we investigate
(Supplement F).

@@

```{r figDiffs, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/diffscmult6ay1aSf1.png"))
file.rename(here::here("analysis/paper/figs_submit/diffscmult6ay1aSf1.pdf"), 
            here::here("analysis/paper/figs_submit/Figure4_diffscmult6ay1aSf1.pdf"))
```

>>>>

resultScenario = "S10"
resultSet ="cmult6ay1aSf1"
figSet = c("diffslow","distScns","EVSI","power","EVPIall")

@@

Examining the probability of correctly assessing population status (Figure \@ref(fig:figStatusErrors))
and the expected value of information (Figures \@ref(fig:figEVSI) and E.1) helps clarify how the need for monitoring varies with landscape condition.
When anthropogenic disturbance is high (right panels of Figures
\@ref(fig:figStatusErrors), \@ref(fig:figEVSI), and E.1)), projections
of population status informed only by priors from the national
demographic-disturbance relationships are accurate, and adding
additional local monitoring data does not improve accuracy.
Interestingly, when the projection period is long and anthropogenic
disturbance is high, we see a negative EVSI (lower right panel of Figure
\@ref(fig:figEVSI)), but we do not see an associated decrease in the
probability of correctly assessing population status (lower right panel
of Figure \@ref(fig:figStatusErrors)). This means that the long-term
projections informed by local monitoring data are less
precise but not less accurate, which can happen if the prior viability predictions are more certain than the posterior predictions (e.g left side of right panels in Figure \@ref(fig:examplePredictions)). At
lower levels of disturbance, local monitoring can increase accuracy, but
there are diminishing returns as the number of years of monitoring increases (Figures
\@ref(fig:figStatusErrors) and \@ref(fig:figEVSI)). Increasing accuracy is more difficult when interannual variation is high (Figures \@ref(fig:figStatusErrors) and \@ref(fig:figEVSI)).The shapes of the
relationships between monitoring effort and the probability of correct
status assessment depend on how anthropogenic disturbance is changing
over time, but there are diminishing benefits of increased monitoring
regardless of whether disturbance is increasing or decreasing
(Supplement D). Acknowledging the possibility of composition survey errors increases uncertainty, but these increases are not large for the set of parameters we investigate
(Supplement F).

Averaging over the full set of example trajectories for each scenario (Figures \@ref(fig:figStatusErrors) and \@ref(fig:figEVSI)) obscures variation among them that is important for monitoring decisions. For example, where disturbance is low but increasing, in the near-term monitoring is most valuable for identifying populations with lower than expected population growth rates (row 1 of Figure \@(fig:figStatusErrorsLow)), but refining knowledge of the current population state also helps identify populations that are likely to remain viable as disturbance increases (row 3 of Figure \@(fig:figStatusErrorsLow)). Status assessment accuracy is much higher among examples where the 95% credible interval for expected population growth rate does not include one (xx %) than among examples where it does (xx %), and this information could be used to guide monitoring. A more nuanced option is to recognize that the posterior probability that the population is viable ($P(\psi|x,\tau)$ in eq \@ref(eq:EVSI)) is an indicator of error probability (Figure \@ref(fig:figProbViability)). One could inform a decision about whether more information is needed for a particular population by selecting an error tolerance (e.g. the probability of correct status assessment should be >0.98) and using that threshold to distinguish cases where the posterior probability of viability is sufficiently certain (i.e. near enough to 0 or 1) from cases where more information is needed (Figure \@ref(fig:figProbViability)).

@@

```{r figDiffs, fig.cap=".", cache=TRUE}
knitr::include_graphics(here::here("figs/forPaper/diffslowcmult6ay1aSf1.png"))
file.rename(here::here("analysis/paper/figs_submit/diffslowcmult6ay1aSf1.pdf"), 
            here::here("analysis/paper/figs_submit/Figure4_diffslowcmult6ay1aSf1.pdf"))
```
